---
title: "Prediction Assignment Writeup"
author: "Alexis Parker"
date: "2023-01-05"
output: html_document
---

### Introduction

This project is using data from 6 participants to predict the way they did an exercise with the use of machine learning.

### Load libraries

```{r}
library(ggplot2)
library(caret)
library(randomForest)
library(rpart)
```

### Load data

```{r}
train <- read.csv("pml-training.csv")
```

### Look at the values

```{r, results='hide'}
head(train)
tail(train)
```

There are two null values: NA and #DIV/0!. There a few columns that should be removed when the data is cleaned.

### Reload and clean the data

```{r}
set.seed(123)

# Loading
train <- read.csv("pml-training.csv", na.strings = c("NA","#DIV/0!",""))
test <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!",""))
```

#### Partition for training data separated into two sections: train and test

```{r}
# Create a partition
in_Train <- createDataPartition(train$classe, p = 0.70, list = FALSE)
train_set <- train[in_Train, ]
test_set <- train[-in_Train, ]
```

```{r}
dim(train_set)
dim(test_set)
```

```{r}
train_set <- train_set[, -c(1:7)]
test_set <- test_set[, -c(1:7)]

train_set <- train_set[, colSums(is.na(train_set)) == 0]
test_set <- test_set[, colSums(is.na(test_set)) == 0]
```

```{r}
dim(train_set)
dim(test_set)
```

## Prediction Models

Random Forest

```{r}
RFmod <- randomForest(as.factor(classe) ~., data = train_set, method = 'class')
RFpreds <- predict(RFmod, newdata = test_set, type = 'class')
# Plot confusion matrix
confusionMatrix(RFpreds, as.factor(test_set$classe))
```

Decision Tree

```{r}
DTmod <- rpart(as.factor(classe) ~., data = train_set, method = 'class')
DTpreds <- predict(DTmod, newdata = test_set, type = 'class')
confusionMatrix(DTpreds, as.factor(test_set$classe))
```

Out of the two models, Random Forest had an higher accuracy, 99.49% than the Decision Tree model, 75.73%. For the final prediction, Random Forest will be used.

### Final prediction

```{r}
# Load and clean the test data for the last prediction
test <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!",""))
test <- test[, -c(1:7)]
test <- test[, colSums(is.na(test)) == 0]

# Prediction
FinalPreds <- predict(RFmod, test)
FinalPreds
```
